---
title: "Weather for Vale de Cavalos, Portugal"
author:
- João Vasco Silva \& Jorge Silva
date: "`r format(Sys.time(), '%d-%B-%Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
    fig_caption: yes
    css: style.css
    header-includes:
    - \usepackage{setspace}
    - \sinlgespacing
---

<body style="background-color:rgb(249, 247, 245);">

```{r, echo=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, echo=FALSE, warning=FALSE}
# package names
packages <- c("leaflet", "plotly", "highcharter", "visNetwork", "DT", "dygraphs", "xts", "htmltools", "tidyverse", "manipulateWidget", "downloadthis")

# install packages
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)){
  install.packages(packages[!installed_packages])}

# load packages
invisible(lapply(packages, library, character.only = TRUE))
```

<hr style="border:1.5px solid black">
# **Background information**

This notebook summarizes time series weather data, starting in 2009 until present, from an automated weather station located in the village of Vale de Cavalos, Santarem, Portugal. The weather station contains sensors to measure solar radiation, minimum and maximum temperature, relative humidity, and precipitation (among others). Reference evapotranspiration was estimated for the location using the Penman-Monteith equations. The weather station maintained by Jorge Silva (please contact jorgesilva.vc60@gmail.com or jvasco323@posteo.net for further details).

The weather data available through this weather station is currently being used to (1) inform decision support tools for disease management in vineyards, (2) support the estimation of crops evolution based on the degree days calculation, and (3) monitor the occurrence of extreme weather events in the region. The same data could be further used to estimate water-limited production yields for the main crops cultivated in the region with crop growth models, among other applications in the field of agronomy. This notebook is updated yearly as data become available.

<hr style="border:1.5px solid black">
# **Location of weather station**

The weather station is installed in a lowland area with a vineyard currently under production, adjacent to the village of Vale de Cavalos (GPS coordinates: 39.288 N, -8.522 W, 250 m above sea level) and approximately 5 km south of Tejo river. The vineyard is part of a transitional zone between human managed heath landscapes ('charneca') and alluvial plains with intensive arable farming ('leziria'). The transitional zone can be best visualized by zooming-out from the location of the station displayed in the map below. The alluvial plains around river Tejo are an important agricultural region for the production of arable crops (e.g., maize and tomato), field vegetables and industrial crops (e.g., tomato and peas), and wine grapes in Portugal. 

```{r, echo=FALSE, fig.align="center", fig.height=2.75, out.width="100%"}
location <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos.csv')
leaflet(location) %>% 
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addCircleMarkers()
```

The region is characterized by a Hot-summer Mediterranean climate (CSa), according the Köppen climate classification (https://en.wikipedia.org/wiki/K%C3%B6ppen_climate_classification). A Mediterranean climate, also called dry summer temperate climate Cs, is a temperate climate sub-type, characterized by dry, hot summers and mild, wet winters. The climate receives its name from the Mediterranean Basin, where this climate type is most common. Mediterranean climate zones are typically located along the western coasts of continents, between roughly 30 and 45 degrees north and south of the equator. The main cause of Mediterranean, or dry summer climate, is the subtropical ridge which extends toward that hemisphere's pole during the summer and migrates toward the equator during the winter due to the seasonal poleward-equatorward variations of temperatures. Please see https://en.wikipedia.org/wiki/Mediterranean_climate#Hot-summer_Mediterranean_climate for further information about the Hot-summer Mediterranean climate.

<hr style="border:1.5px solid black">
# **Daily weather conditions**

```{r, echo=FALSE, fig.height=10, fig.asp=0.618, out.width="100%", fig.align="center"}

date_start <- "2024-01-01"
date_end <- "2024-12-31"

# plot 1
data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')
data_aggr <- aggregate(data[c('IRRAD_MJm2day', 'UM')], by=list('year'=data$Year, 'month'=data$Month), FUN=mean, na.rm=T)
ts_data_rad <- xts(data[,47], order.by=as.Date(data[,1], "%Y-%m-%d"))
ts_data_hum <- xts(data[,9], order.by=as.Date(data[,1], "%Y-%m-%d"))
ts_rad <- cbind(ts_data_rad, ts_data_hum)
rad <- dygraph(ts_rad, group=ts_rad) %>%
  dyAxis("y", label = "Solar radiation (MJ/m2/day)", valueRange = c(0, 55), independentTicks=TRUE, axisLabelWidth=35)%>%
  dyAxis("y2", label = "Relative humidity (%)", valueRange = c(0, 110), independentTicks=TRUE, axisLabelWidth=35) %>%
  dySeries("ts_data_rad", label='Radiation', color='goldenrod', axis='y') %>%
  dySeries("ts_data_hum", label='Humidity', color='steelblue', axis='y2') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2) %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE)

# plot 2
data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')
data_aggr <- aggregate(data[c('TMIN', 'TMAX')], by=list('year'=data$Year, 'month'=data$Month), FUN=mean, na.rm=T)
ts_data_tmin <- xts(data[,2], order.by=as.Date(data[,1], "%Y-%m-%d"))
ts_data_tmax <- xts(data[,3], order.by=as.Date(data[,1], "%Y-%m-%d"))
ts_temp <- cbind(ts_data_tmin, ts_data_tmax)
temp <- dygraph(ts_temp, group=ts_rad) %>%
  dyAxis("y", label='Air temperature (C)', valueRange = c(-10, 55), axisLabelWidth=35) %>% 
  dyAxis("y2", label='Air temperature (C)', valueRange = c(-10, 55), axisLabelWidth=35) %>% 
  dySeries("ts_data_tmax", label="TMAX", axis='y') %>%
  dySeries("ts_data_tmin", label="TMIN", axis='y2') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2) %>%
  dyLimit(40, "", strokePattern="solid", color="red") %>% 
  dyLimit(0, "", strokePattern="solid", color="red") %>% 
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE)

# plot 3
data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')
data_aggr <- aggregate(data[c('RAIN_mm', 'ET0_mm')], by=list('year'=data$Year, 'month'=data$Month), FUN=mean, na.rm=T)
ts_data_rain <- xts(data[,15], order.by=as.Date(data[,1], "%Y-%m-%d"))
ts_data_et0 <- xts(data[,26], order.by=as.Date(data[,1], "%Y-%m-%d"))
ts_rain <- cbind(ts_data_rain, ts_data_et0)
rain <- dygraph(ts_rain, group=ts_rad) %>%
  dyAxis("y", label = "Precipitation (mm)", valueRange = c(0, 65), independentTicks = TRUE, axisLabelWidth=35)%>%
  dyAxis("y2", label = "Reference ET (mm)", valueRange = c(0, 6.5), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("ts_data_rain", label='RAIN', color='darkblue', axis='y') %>%
  dySeries("ts_data_et0", label='ET0', color='dodgerblue', axis='y2') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2) %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyRangeSelector(height = 40, dateWindow = c(date_start, date_end))

# both
combineWidgets(rad, temp, rain, nrow=3, ncol=1, rowsize=c(1,1,1), height=720)
```

<hr style="border:1.5px solid black">
# **Climate change over time**

```{r, echo=FALSE, fig.height=10, fig.asp=0.3, out.width="100%", fig.align="center"}

# radiation & rain -------------------------------------------------------------

# load data
data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')

# january - december
data_yr <- data
data_yr <- aggregate(data_yr[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_yr$Year), FUN=sum, na.rm=T)
data_yr$year <- paste0(data_yr$year, '-01-01')
data_yr_rain_cum <- xts(data_yr[,2], order.by=as.Date(data_yr$year))  
data_yr_srad_cum <- xts(data_yr[,3], order.by=as.Date(data_yr$year))  

# january - march
data_mar <- subset(data, Month == 1 | Month == 2 | Month == 3)
data_mar <- aggregate(data_mar[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_mar$Year), FUN=sum, na.rm=T)
data_mar$year <- paste0(data_mar$year, '-01-01')
data_mar_rain_cum <- xts(data_mar[,2], order.by=as.Date(data_mar$year))  
data_mar_srad_cum <- xts(data_mar[,3], order.by=as.Date(data_mar$year))  

# january - june
data_jun <- subset(data, Month == 1 | Month == 2 | Month == 3 | Month == 4 | Month == 5 | Month == 6)
data_jun <- aggregate(data_jun[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_jun$Year), FUN=sum, na.rm=T)
data_jun$year <- paste0(data_jun$year, '-01-01')
data_jun_rain_cum <- xts(data_jun[,2], order.by=as.Date(data_jun$year))  
data_jun_srad_cum <- xts(data_jun[,3], order.by=as.Date(data_jun$year))  

# january - september
data_sep <- subset(data, Month == 1 | Month == 2 | Month == 3 | Month == 4 | Month == 5 | Month == 6 | Month == 7 | Month == 8 | Month == 9)
data_sep <- aggregate(data_sep[c('RAIN_mm', 'IRRAD_MJm2day')], by=list('year'=data_sep$Year), FUN=sum, na.rm=T)
data_sep$year <- paste0(data_sep$year, '-01-01')
data_sep_rain_cum <- xts(data_sep[,2], order.by=as.Date(data_sep$year))  
data_sep_srad_cum <- xts(data_sep[,3], order.by=as.Date(data_sep$year))  

# data for plot
data_plot_srad <- cbind(data_mar_srad_cum/100, data_jun_srad_cum/100, data_sep_srad_cum/100, data_yr_srad_cum/100)
p1 <- dygraph(data_plot_srad) %>%
  dyAxis("y", label = "Solar radiation (x100 MJ/m2)", valueRange = c(0, 80), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_srad_cum.100", label='Year', color='orangered', axis='y') %>%
  dySeries("data_mar_srad_cum.100", label='Jan-Mar', color='goldenrod', axis='y') %>%
  dySeries("data_jun_srad_cum.100", label='Apr-Jun', color='goldenrod', axis='y') %>%
  dySeries("data_sep_srad_cum.100", label='Jul-Sep', color='goldenrod', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# data for plot
data_plot_rain <- cbind(data_mar_rain_cum/10, data_jun_rain_cum/10, data_sep_rain_cum/10, data_yr_rain_cum/10)
p2 <- dygraph(data_plot_rain, group=data_plot_srad) %>%
  dyAxis("y", label = "Precipitation (cm)", valueRange = c(0, 110), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_rain_cum.10", label='Year', color='darkblue', axis='y') %>%
  dySeries("data_mar_rain_cum.10", label='Jan-Mar', color='dodgerblue', axis='y') %>%
  dySeries("data_jun_rain_cum.10", label='Apr-Jun', color='dodgerblue', axis='y') %>%
  dySeries("data_sep_rain_cum.10", label='Jul-Sep', color='dodgerblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") 

# temperature ------------------------------------------------------------------

# january - december
data_yr <- data
data_yr <- aggregate(data_yr[c('TMAX', 'TMIN')], by=list('year'=data_yr$Year), FUN=mean, na.rm=T)
data_yr$year <- paste0(data_yr$year, '-01-01')
data_yr_tmax_av <- xts(data_yr[,2], order.by=as.Date(data_yr$year))  
data_yr_tmin_av <- xts(data_yr[,3], order.by=as.Date(data_yr$year))  

# january - march
data_mar <- subset(data, Month == 1 | Month == 2 | Month == 3)
data_mar <- aggregate(data_mar[c('TMAX', 'TMIN')], by=list('year'=data_mar$Year), FUN=mean, na.rm=T)
data_mar$year <- paste0(data_mar$year, '-01-01')
data_mar_tmax_av <- xts(data_mar[,2], order.by=as.Date(data_mar$year))  
data_mar_tmin_av <- xts(data_mar[,3], order.by=as.Date(data_mar$year))  

# april - june
data_jun <- subset(data, Month == 4 | Month == 5 | Month == 6)
data_jun <- aggregate(data_jun[c('TMAX', 'TMIN')], by=list('year'=data_jun$Year), FUN=mean, na.rm=T)
data_jun$year <- paste0(data_jun$year, '-01-01')
data_jun_tmax_av <- xts(data_jun[,2], order.by=as.Date(data_jun$year))  
data_jun_tmin_av <- xts(data_jun[,3], order.by=as.Date(data_jun$year))  

# july - september
data_sep <- subset(data, Month == 7 | Month == 8 | Month == 9)
data_sep <- aggregate(data_sep[c('TMAX', 'TMIN')], by=list('year'=data_sep$Year), FUN=mean, na.rm=T)
data_sep$year <- paste0(data_sep$year, '-01-01')
data_sep_tmax_av <- xts(data_sep[,2], order.by=as.Date(data_sep$year))  
data_sep_tmin_av <- xts(data_sep[,3], order.by=as.Date(data_sep$year))  

# october - december
data_dec <- subset(data, Month == 10 | Month == 11 | Month == 12)
data_dec <- aggregate(data_dec[c('TMAX', 'TMIN')], by=list('year'=data_dec$Year), FUN=mean, na.rm=T)
data_dec$year <- paste0(data_dec$year, '-01-01')
data_dec_tmax_av <- xts(data_dec[,2], order.by=as.Date(data_dec$year))  
data_dec_tmin_av <- xts(data_dec[,3], order.by=as.Date(data_dec$year))  

# data for plot
data_plot_tmax <- cbind(data_mar_tmax_av, data_jun_tmax_av, data_sep_tmax_av, data_dec_tmax_av, data_yr_tmax_av)
p3 <- dygraph(data_plot_tmax) %>%
  dyAxis("y", label = "Maximum temperature (C)", valueRange = c(10, 45), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_tmax_av", label='Year', color='black', axis='y') %>%
  dySeries("data_mar_tmax_av", label='Jan-Mar', color='red', axis='y') %>%
  dySeries("data_jun_tmax_av", label='Apr-Jun', color='salmon', axis='y') %>%
  dySeries("data_sep_tmax_av", label='Jul-Sep', color='plum', axis='y') %>%
  dySeries("data_dec_tmax_av", label='Oct-Dec', color='firebrick', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# data for plot
data_plot_tmin <- cbind(data_mar_tmin_av, data_jun_tmin_av, data_sep_tmin_av, data_dec_tmin_av, data_yr_tmin_av)
p4 <- dygraph(data_plot_tmin) %>%
  dyAxis("y", label = "Minimum temperature (C)", valueRange = c(-2, 20), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("data_yr_tmin_av", label='Year', color='black', axis='y') %>%
  dySeries("data_mar_tmin_av", label='Jan-Mar', color='blue', axis='y') %>%
  dySeries("data_jun_tmin_av", label='Apr-Jun', color='steelblue', axis='y') %>%
  dySeries("data_sep_tmin_av", label='Jul-Sep', color='skyblue', axis='y') %>%
  dySeries("data_dec_tmin_av", label='Oct-Dec', color='royalblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# number of days ---------------------------------------------------------------

data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')
heat_days <- subset(data, TMAX > 39)
heat_days$unique <- 1
heat_days <- aggregate(heat_days[c('unique')], by=list('year'=heat_days$Year), FUN=sum, na.rm=T)
heat_days$year <- paste0(heat_days$year, '-01-01')
heat_days <- xts(heat_days[,2], order.by=as.Date(heat_days$year))  
cold_days <- subset(data, TMIN < 0)
cold_days$unique <- 1
cold_days <- aggregate(cold_days[c('unique')], by=list('year'=cold_days$Year), FUN=sum, na.rm=T)
cold_days$year <- paste0(cold_days$year, '-01-01')
cold_days <- xts(cold_days[,2], order.by=as.Date(cold_days$year))  

# data for plot
data_plot_days <- cbind(heat_days, cold_days)
p5 <- dygraph(data_plot_days) %>%
  dyAxis("y", label = "Number of heat & cold days", valueRange = c(0, 60), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("heat_days", label='Heat days (TMAX > 39)', color='firebrick', axis='y') %>%
  dySeries("cold_days", label='Cold days (TMIN < 0)', color='royalblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }") # %>%

# number of days ---------------------------------------------------------------

rain_days <- subset(data, RAIN_mm > 0)
rain_days$unique <- 1
rain_days <- aggregate(rain_days[c('unique')], by=list('year'=rain_days$Year), FUN=sum, na.rm=T)
rain_days$year <- paste0(rain_days$year, '-01-01')
rain_days <- xts(rain_days[,2], order.by=as.Date(rain_days$year))  

# data for plot
data_plot_days <- cbind(heat_days, cold_days, rain_days)[,c(3)]
p6 <- dygraph(data_plot_days) %>%
  dyAxis("y", label = "Number of rain days", valueRange = c(60, 160), independentTicks = TRUE, axisLabelWidth=35) %>%
  dySeries("rain_days", label='Rain days (Rain > 0)', color='royalblue', axis='y') %>%
  dyOptions(fillGraph=T, drawPoints=T, pointSize=2, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("x", axisLabelFormatter="function(d) { return d.getFullYear() }")

# all plots --------------------------------------------------------------------

# both
combineWidgets(p1, p2, p3, p4, p5, p6, nrow=3, ncol=2, rowsize=c(1), height=550+275)

#plotInteractive <- function(Months, Variable){
#  # subset
#  month_subset <- subset(data, Month %in% Months)
#  data_yr <- aggregate(month_subset[c(Variable)], by=list('year'=month_subset$Year), FUN=mean, na.rm=T)
#  # plot
#  dygraph(data_yr, main = Variable) %>%
#    dyAxis("y", label = Variable) %>% #, valueRange = c(12, 18)) %>%
#    dySeries(Variable, label=Variable, axis='y') %>%
#    dyOptions(axisLineWidth = 1.5, fillGraph = TRUE, drawGrid = TRUE, drawPoints=T, pointSize=2) %>%
#    dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) }
# manipulateWidget(
#  plotInteractive(Months, Variable),
#  Variable = mwSelect(sort(unique(c('IRRAD_MJm2day', 'RAIN_mm'))), "IRRAD_MJm2day"),
#  Months = mwCheckboxGroup(unique(data$Month), 0)
# )

```

<hr style="border:1.5px solid black">
# **Cumulative values**

```{r, echo=FALSE, fig.height=10, fig.asp=0.3, out.width="100%", fig.align="center"}

data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')

# SRAD
srad <- data[c('Y_year', 'DOY', 'IRRAD_MJm2day')]
srad_final <- c()
for(yr in unique(srad$Y_year)){
  srad_yr <- subset(srad, Y_year == yr)
  srad_yr$IRRAD_MJm2day <- cumsum(srad_yr$IRRAD_MJm2day)
  srad_final <- rbind(srad_final, srad_yr)}
srad_final$IRRAD_MJm2day <- srad_final$IRRAD_MJm2day/100
srad <- reshape2::dcast(srad_final, DOY ~ Y_year, value.var='IRRAD_MJm2day')
cum_srad <- dygraph(srad) %>%
  dyGroup(c(names(srad)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::inferno(14)) %>%
  dyAxis("y", label = "Solar radiation (x100 MJ/m2/d)", valueRange = c(0, 80), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# GDD
gdd <- data[c('Y_year', 'DOY', 'GDD')]
gdd_final <- c()
for(yr in unique(gdd$Y_year)){
  gdd_yr <- subset(gdd, Y_year == yr)
  gdd_yr$GDD <- cumsum(gdd_yr$GDD)
  gdd_final <- rbind(gdd_final, gdd_yr)}
gdd_final$GDD <- gdd_final$GDD/100
gdd <- reshape2::dcast(gdd_final, DOY ~ Y_year, value.var='GDD')
gdd_val <- dygraph(gdd) %>%
  dyGroup(c(names(gdd)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::inferno(14)) %>%
  dyAxis("y", label = "Growing degrees day (x100)", valueRange = c(0, 35), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# RAIN
rain <- data[c('Y_year', 'DOY', 'RAIN_mm')]
rain_final <- c()
for(yr in unique(rain$Y_year)){
  rain_yr <- subset(rain, Y_year == yr)
  rain_yr$RAIN_mm <- cumsum(rain_yr$RAIN_mm)
  rain_final <- rbind(rain_final, rain_yr)}
rain_final$RAIN_mm <- rain_final$RAIN_mm/10
rain <- reshape2::dcast(rain_final, DOY ~ Y_year, value.var='RAIN_mm')
cum_rain <- dygraph(rain) %>%
  dyGroup(c(names(rain)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::viridis(14)) %>%
  dyAxis("y", label = "Precipitation (cm)", valueRange = c(0, 130), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# ET
et <- data[c('Y_year', 'DOY', 'ET0_mm')]
et_final <- c()
for(yr in unique(et$Y_year)){
  et_yr <- subset(et, Y_year == yr)
  et_yr$ET0_mm <- cumsum(et_yr$ET0_mm)
  et_final <- rbind(et_final, et_yr)}
et_final$ET0_mm <- et_final$ET0_mm/10
et <- reshape2::dcast(et_final, DOY ~ Y_year, value.var='ET0_mm')
cum_et <- dygraph(et) %>%
  dyGroup(c(names(et)[-1]), stepPlot = TRUE, drawPoints = TRUE, color = viridis::viridis(14)) %>%
  dyAxis("y", label = "Reference ET (cm)", valueRange = c(0, 140), independentTicks = TRUE, axisLabelWidth=35) %>%
  dyOptions(fillGraph=F, drawPoints=T, pointSize=1, drawYAxis=T, drawXAxis=T, axisLineColor='black') %>%
  dyHighlight(highlightCircleSize=4, highlightSeriesBackgroundAlpha = 0.5, highlightSeriesOpts = list(), hideOnMouseOut = TRUE) %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyLegend(width=300)

# pal <- colorRampPalette(c('darkred', "darkorange", 'orange', 'yellowgreen', 'forestgreen', 'darkgreen'))
# cum_et <- hchart(et_final, "line", hcaes(x=DOY, y=ET0_mm, group=Year, color=Year)) %>%
#   hc_colors(c(pal(length(unique(et_final$Year))))) %>%
#   hc_legend() %>%
#   hc_yAxis(title = list(text="Reference ET (mm)", style=list(color='black', fontSize=16)),
#            labels = list(style=list(color='black', fontSize=15))) %>%  
#   hc_xAxis(title = list(text="Day-of-the-year (DOY)", style=list(color='black', fontSize=16)), 
#            labels = list(style=list(color='black', fontSize=15)),  
#            tickInterval = 1) %>%
#   hc_chart(events = list(load = JS("function() {
#       var chart = this; 
#       chart.update({ chart: { plotBorderColor: 'black', plotBorderWidth: '1' } } ); }")))

# both
combineWidgets(cum_rain, gdd_val, cum_et, cum_srad, nrow=2, ncol=2, rowsize=c(1), height=600+100) 

```

<hr style="border:1.5px solid black">
# **Table with complete data**

```{r, echo=FALSE, warning=FALSE}
data <- read.csv('D:/# Jvasco/Visualization/Vale de Cavalos/rmarkdown-weather/data/vale_de_cavalos_weather.csv')
data <- data[c('Station', 'Date', 'Day', 'Month', 'Year', 'DOY', 'IRRAD_MJm2day', 'UM', 
               'TMAX', 'TMIN', 'GDD', 'RAIN_mm', 'ET0_mm', 'LAT', 'LONG', 'ELEV')]
colnames(data)[7] <- 'SRAD'
round_df <- function(x, digits) {
    numeric_columns <- sapply(x, mode) == 'numeric'
    x[numeric_columns] <-  round(x[numeric_columns], digits)
    x}
data <- round_df(data, 2)
data %>%
  DT::datatable(options=list(scrollX=TRUE, scrollCollapse=TRUE), rownames = FALSE)
```

```{r, echo=FALSE, warning=FALSE}
data %>%
   download_this(
     output_name = "weather_data_vale_de_cavalos",
     output_extension = ".csv",
     button_label = "DOWNLOAD DATA AS CSV",
     button_type = "default",
     has_icon = TRUE,
     icon = "fa fa-save")
```
